Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2016-01-20 latex/article.tex
Added figures

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\color{blue}
Here, we focus on curvature tensor estimation (mean/Gaussian curvature, principal curvatures
directions, \ldots). Most curvature estimators require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface, 
and is naturally related to the amount of perturbation on input
data\todo{DC: ``related to a expected signal/noise ratio?''}. As a consequence, when using
such estimators, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
needs\todo{DC: needs->need?} to be done.
\normalcolor
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

@@

\vspace{0.2cm} \color{blue} In this paper, we first present the
previous works on curvature tensor estimation. Then, we propose na
efficient GPU approach to extract a triangulated isosurface from a
digital object. Finally, we propose a fully-parallel  GPU pipeline to compute
curvature tensor in real time and discuss on experiments. \todo{DC:
  bof le discuss}
\normalcolor

@@

\txtblue{To implement the integration on $\Ball{R}{\vx} \cap \Shape$
  (Fig.~\ref{fig:approx}-$a$), several strategies have been evaluated.}

@@

\paragraph{Naive Approach.}
\txtblue{A first simple solution  consists in scanning all digital
points, at a given resolution, lying inside the integration domain}
(Fig.~\ref{fig:approx}-$b$) and then \txtblue{estimating} the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection (see
Eq.~(\ref{eq:dig-curvature-estimator-k}) to (\ref{eq:dig-curvature-estimator-k2})). 
\color{blue}
This is exactly similar to what is done on the CPU. On  the GPU, we can
exploit \emph{mipmap} textures to obtain  multi-resolution information.
\normalcolor
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\color{blue}
a given level $l$, a texture probe at  a point $(x,y,z)$ returns the fraction of $\DSh$ belonging
\normalcolor
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$c,d$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

\begin{figure}
  \vspace{-0.35cm}
  \begin{center}
    {\includegraphics[width=4.0cm]{figs/approx1}}
    {\includegraphics[width=4.0cm]{figs/approx-reg-1}}
    {\includegraphics[width=4.0cm]{figs/approx-reg-2}}
    %TODO image Mipmap 3eme Niveau
  \end{center}
  \vspace{-0.35cm}
  \caption{Integral computation with an area estimation (\emph{left}) 
  \color{blue}
  	from mipmaps at level $l=0$ (note that this is the same as pixel counting) \emph{middle-left}
  	from mipmaps at level $l=1$ \emph{middle-right}
  	from mipmaps at level $l=2$ \emph{right}
  	We can note that the higher the mipmap level, the more we approximate our integrand, and the less texture probe we need to do.
    \normalcolor}
  \label{fig:approx}
  \vspace{-0.35cm}
\end{figure}

\begin{figure}
  \vspace{-0.35cm}
  \begin{center}
    {\includegraphics[width=4cm]{figs/approx-bh}}
    %{\includegraphics[width=4cm]{figs/approx-ada}}
    %TODO Image alignement cellules mipmaps
  \end{center}
  \vspace{-0.35cm}
  \caption{Adaptive approximation of the integration using a
    hierarchical decomposition of the ball $\Ball{R}{\vx}$ }
    %(\emph{left}), or a decomposition of $\Ball{R}{\vx}\cap\Shape$ (\emph{right}).}
  \label{fig:approx2}
  \vspace{-0.35cm}
\end{figure}

@@

\paragraph{Hierarchical Decomposition.}  
\color{blue}
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$
(Fig.~\ref{fig:approx2}-$a$) 
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$). 
The idea is to decompose the
ball into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. 
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$. 
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$.
\txtred{On the GPU, computing a hierarchical decomposition of
  $B_R(\vx)$ at each point $\vx$ is highly inefficient since the
  hardware optimizes the parallelism only when the micro-program
  described in the shader has a predictive execution flow. Hence,
  implementing the recursive algorithm (or its de-recursified version)
  requires a lot of branches (conditional \emph{if} tests) in the
  flow.}  We have thus considered a fast multi-resolution
approximation as illustrated in Fig.~\ref{fig:approx2}-$a$: for a
given radius $R$, we precompute a hierarchical decomposition of
$B_R$. Such decomposition is made of octree cells at different
resolutions. At a given MC vertex $\vx$, the algorithm becomes simple
since we just scan the ball cells and estimate the area (or other
moments) from the mipmap values associated to this cell. Note that
these \txtred{precomputed} cells in the $B_R$ decomposition may not be aligned
with mipmap cells. However, we can use GPU %mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells in
$B_R$ is in $O(\log{R})$. Implementation details on the hierarchical
octree representation can be found in the supplementary material (see
Section \ref{sec:discussion}). \sout{ Still, this method is not fully
exploiting the GPU parallelism.  The hardware optimizes the
parallelism only when the micro-program described in the shader has a
predictive execution flow. Hence, implementing this recursive
algorithm (or its de-recursified version) requires a lot of branches
(conditional \emph{if} tests) in the flow.}\todo{la discussion vient apres}  \normalcolor

\paragraph{Dynamic Refinement.} In this last approach, we want to optimize
the interactivity and the execution flow or parallelism on GPU. The
integration is simply computed using a regular grid at different
resolution $l$ (Fig. \ref{fig:approx}-$b$ and $c$). The shader code
becomes trivial (simple triple loop) and a lot of texture fetches are
involved, but interactivity can be easily obtained. Indeed, we
consider the following multi-pass approach:
\begin{enumerate}
\item When the surface geometry has been computed, we set the current
  level $l$ at an high level $l\EqDef l_{max}$ of the mipmap
  pyramid.
\item We compute the integrals (and the curvature tensor) at the
  mipmap level $l$ and  send the estimated quantities to the
  visualization stage.
\item If the user changes the camera position or the computation
  parameters, we return to step 1.
\item If the current framerate is above a given threshold, we decrease
  $l$ and return to step 2 if the final $l_0$ level has not been
  reached yet.
\end{enumerate}
A consequence of the multi-pass approach is that when there is no
interaction (camera settings, parameters), the GPU automatically
refines the estimation. Even if step 2 is quite expensive, in
$O\left(\left(\frac{R}{2^l}\right)^3\right)$, the interactive control
in step 3 considerably improves the interactive exploration of the
tensor with fast preliminary approximations which are quickly refined.

>>>>

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\color{blue}
Here, we focus on curvature tensor estimation (mean/Gaussian curvature, principal curvatures
directions, \ldots). Most curvature estimators require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface, 
and is naturally related to the amount of perturbation on input
data\todo{DC: ``related to a expected signal/noise ratio?''}. As a consequence, when using
such estimators, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
need to be done.
\normalcolor
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

@@

\vspace{0.2cm} \color{blue} In this paper, we first present the
previous works on curvature tensor estimation. Then, we propose na
efficient GPU approach to extract a triangulated isosurface from a
digital object. Finally, we propose a fully-parallel  GPU pipeline to compute
curvature tensor in real time and present our results.
\normalcolor

@@

\txtblue{To implement the integration on $\Ball{R}{\vx} \cap \Shape$
  (Fig.~\ref{fig:notations}), several strategies have been evaluated.}

@@

\paragraph{Naive Approach.}
\txtblue{A first simple solution  consists in scanning all digital
points, at a given resolution, lying inside the integration domain}
(Fig.~\ref{fig:approx}-$left$) and then \txtblue{estimating} the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection (see
Eq.~(\ref{eq:dig-curvature-estimator-k}) to (\ref{eq:dig-curvature-estimator-k2})). 
\color{blue}
This is exactly similar to what is done on the CPU. On  the GPU, we can
exploit \emph{mipmap} textures to obtain  multi-resolution information.
\normalcolor
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\color{blue}
a given level $l$, a texture probe at  a point $(x,y,z)$ returns the fraction of $\DSh$ belonging
\normalcolor
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$right$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

\begin{figure}
  \vspace{-0.35cm}
  \begin{center}
    {\includegraphics[width=4.0cm]{figs/approx-naive}}
    {\includegraphics[width=4.0cm]{figs/approx-hierarchical}}
    {\includegraphics[width=4.0cm]{figs/approx-dynamic}}
  \end{center}
  \vspace{-0.35cm}
  \caption{\txtblue{
  	Here are the three approaches we used to evaluate the integrand.
  	(\emph{left}) is the naive approach, where we sum all the digital points that are inside the integrand domain
  	(\emph{middle}) shows the hierarchical decomposition of the ball, thus limiting the number of texture probes to do.
  	(\emph{right}) illustrates the dynamic approach, we probe the textures at a higher resolution and we refine it at each frame until we reach the finer level.
    }}
  \label{fig:approx}
  \vspace{-0.35cm}
\end{figure}

%\begin{figure}
%  \vspace{-0.35cm}
%  \begin{center}
%    {\includegraphics[width=4cm]{figs/approx-bh}}
%    %{\includegraphics[width=4cm]{figs/approx-ada}}
%  \end{center}
%  \vspace{-0.35cm}
%  \caption{Adaptive approximation of the integration using a
%    hierarchical decomposition of the ball $\Ball{R}{\vx}$ }
%    %(\emph{left}), or a decomposition of $\Ball{R}{\vx}\cap\Shape$ (\emph{right}).}
%  \label{fig:approx2}
%  \vspace{-0.35cm}
%\end{figure}

@@

\paragraph{Hierarchical Decomposition.}  
\color{blue}
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$.
%(Fig.~\ref{fig:approx}-$middle$) 
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$). 
The idea is to decompose the
ball into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. 
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$. 
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$.
\txtred{On the GPU, computing a hierarchical decomposition of
  $B_R(\vx)$ at each point $\vx$ is highly inefficient since the
  hardware optimizes the parallelism only when the micro-program
  described in the shader has a predictive execution flow. Hence,
  implementing the recursive algorithm (or its de-recursified version)
  requires a lot of branches (conditional \emph{if} tests) in the
  flow.}  We have thus considered a fast multi-resolution
approximation as illustrated in Fig.~\ref{fig:approx}-$middle$: for a
given radius $R$, we precompute a hierarchical decomposition of
$B_R$. Such decomposition is made of octree cells at different
resolutions. At a given MC vertex $\vx$, the algorithm becomes simple
since we just scan the ball cells and estimate the area (or other
moments) from the mipmap values associated to this cell. Note that
these \txtred{precomputed} cells in the $B_R$ decomposition may not be aligned
with mipmap cells. However, we can use GPU %mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells in
$B_R$ is in $O(\log{R})$. Implementation details on the hierarchical
octree representation can be found in the supplementary material (see
Section \ref{sec:discussion}). \sout{ Still, this method is not fully
exploiting the GPU parallelism.  The hardware optimizes the
parallelism only when the micro-program described in the shader has a
predictive execution flow. Hence, implementing this recursive
algorithm (or its de-recursified version) requires a lot of branches
(conditional \emph{if} tests) in the flow.}\todo{la discussion vient apres}  \normalcolor

\paragraph{Dynamic Refinement.} In this last approach, we want to optimize
the interactivity and the execution flow or parallelism on GPU. The
integration is simply computed using a regular grid at different
resolution $l$ (Fig. \ref{fig:approx}-$right$). The shader code
becomes trivial (simple triple loop) and a lot of texture fetches are
involved, but interactivity can be easily obtained. Indeed, we
consider the following multi-pass approach:
\begin{enumerate}
\item When the surface geometry has been computed, we set the current
  level $l$ at an high level $l\EqDef l_{max}$ of the mipmap
  pyramid.
\item We compute the integrals (and the curvature tensor) at the
  mipmap level $l$ and  send the estimated quantities to the
  visualization stage.
\item If the user changes the camera position or the computation
  parameters, we return to step 1.
\item If the current framerate is above a given threshold, we decrease
  $l$ and return to step 2 if the final $l_0$ level has not been
  reached yet.
\end{enumerate}
A consequence of the multi-pass approach is that when there is no
interaction (camera settings, parameters), the GPU automatically
refines the estimation. Even if step 2 is quite expensive, in
$O\left(\left(\frac{R}{2^l}\right)^3\right)$, the interactive control
in step 3 considerably improves the interactive exploration of the
tensor with fast preliminary approximations which are quickly refined.
